---
title: "Data preparation"
output:
  pdf_document: default
---

# Instructions

- You only need to submit the .Rmd of this file, not a PDF.

- You should __comment__ your code clearly to show what you've done to prepare the data.

- The purpose of this file is to use the data in the `data-raw` folder to create the data you will use in the report. The data you will use in the report should be saved in the `data` folder. It is good professional practice to make sure you're never directly modifying your raw data, but instead creating new datasets based on merges/manipulations that you need to reuse.

- Make sure you've taken a look at the hints for the web scraping and census API. 

- You may find the `write_rds()` function from the `readr` package helpful (it is loaded as part of the `tidyverse`).

- You do not need to keep the structure below.

# Set up

```{r, libraries}
# Set up any libraries you need
library(tidyverse)
library(polite)
library(rvest)
library(lubridate)
library(haven)
library(cancensus)

```

# Loading client data
```{r}
emergency <- readRDS("data-raw/break_glass_in_case_of_emergency.Rds")
cust_dev <- readRDS("data-raw/cust_dev.Rds")
cust_sleep <- readRDS("data-raw/cust_sleep.Rds")
customer <- readRDS("data-raw/customer.Rds")
device <- readRDS("data-raw/device.Rds")
```

# Getting external data

## Web scraping industry data

```{r}
url <- "https://fitnesstrackerinfohub.netlify.app/"

# Make sure this code is updated appropriately to provide 
# informative user_agent details
target <- bow(url,
              user_agent = "ericyanqi.liu@mail.utoronto.ca for STA303/1002 project",
              force = TRUE)

# Any details provided in the robots text on crawl delays and 
# which agents are allowed to scrape
target

html <- scrape(target)

device_data <- html %>% 
  html_elements("table") %>% 
  html_table() %>% 
  pluck(1) # added, in case you're getting a list format
```

# Postal Code conversion

```{r}
# install.packages("haven")
dataset = readRDS("data-raw/break_glass_in_case_of_emergency.Rds")

postcode <- dataset %>% 
  dplyr::select(PC, CSDuid) %>% 
  rename(c("postcode" = "PC"))
```

# Census API

```{r}



options(cancensus.api_key = "CensusMapper_b20f6d370fdaee5e01db7b4e82791689",
        cancensus.cache_path = "cache") # this sets a folder for your cache


# get all regions as at the 2016 Census (2020 not up yet)
regions <- list_census_regions(dataset = "CA16")

regions_filtered <-  regions %>% 
  filter(level == "CSD") %>% # Figure out what CSD means in Census data
  as_census_region_list()

# This can take a while
# We want to get household median income
census_data_csd <- get_census(dataset='CA16', regions = regions_filtered,
                          vectors=c("v_CA16_2397"), 
                          level='CSD', geo_format = "sf")

# Simplify to only needed variables
median_income <- census_data_csd %>% 
  as_tibble() %>% 
  dplyr::select(CSDuid = GeoUID, contains("median"), Population) %>% 
  mutate(CSDuid = parse_number(CSDuid)) %>% 
  rename(hhld_median_inc = 2)

```

# merging datasets

```{r}
#getting median income and population information for report
median_income_postcode <- postcode %>% 
  distinct_all() %>% 
  left_join(median_income) %>% 
  group_by(postcode) %>% 
  summarise(hhld_median_inc = mean(hhld_median_inc, na.rm = TRUE),
            Population = sum(Population, na.rm = TRUE)) %>% 
  ungroup()

#combing data into one large data set (left join was used)
cust_dev_combine <- customer %>% 
  ungroup() %>% 
  left_join(cust_dev) %>% 
  left_join(device) %>% 
  left_join(rename(device_data, c("device_name" = "Device name", "line" = "Line"))) %>% 
  left_join(median_income_postcode)


#attaining skin tone information based off emoji preference
cust_dev_emoji_mod<- cust_dev_combine %>% 
  mutate(age = as.numeric(2022 - year(dob))) %>% 
  mutate(skin_tone = case_when(is.na(emoji_modifier) ~ "yellow",
                               emoji_modifier == "U+1F3FB" ~ "light",
                               emoji_modifier == "U+1F3FC" ~ "medium-light",
                               emoji_modifier == "U+1F3FD" ~ "medium",
                               emoji_modifier == "U+1F3FE" ~ "medium-dark",
                               emoji_modifier == "U+1F3FF" ~ "dark"))

# separating customers as traditional and non tradional customers
cust_dev4<- cust_dev_emoji_mod %>%
   mutate(custType = ifelse(line == "Active" | line == "Advance", "Non_Traditional", "Traditional"))

cust_dev5<- cust_dev_emoji_mod %>%
   mutate(custTypeNum = ifelse(line == "Active" | line == "Advance", 1, 0))

# Remove missing data observations
cust_dev4<- na.omit(cust_dev4)

cust_dev5<- na.omit(cust_dev5)

# write data to data folder
write_rds(cust_dev4, "data/cust_dev4.Rds")

write_rds(cust_dev5, "data/cust_dev5.Rds")


```




```{r}
# Join cust_dev and device data, and renamed column variables like Device name and Line to make it easier to work with
cust_dev2 <- customer %>% 
  ungroup() %>% 
  left_join(cust_dev) %>% 
  left_join(device) %>% 
  left_join(rename(device_data, c("device_name" = "Device name", "line" = "Line")))

# Final cust_dev dataset
# Create new column variable that labels Active/Advance devices as Non-traditional and the rest as Traditional
# Create another new column variable that translates the emoji modifier into decipherable skin tones
cust_dev3 <- cust_dev2 %>% 
  mutate(Active_OR_Advance = if_else(line %in% c("Active", "Advance"), "Non-traditional", "Traditional")) %>% 
  mutate(skin_tone = case_when(is.na(emoji_modifier) ~ "yellow",
                               emoji_modifier == "U+1F3FB" ~ "light",
                               emoji_modifier == "U+1F3FC" ~ "medium-light",
                               emoji_modifier == "U+1F3FD" ~ "medium",
                               emoji_modifier == "U+1F3FE" ~ "medium-dark",
                               emoji_modifier == "U+1F3FF" ~ "dark")) 
# Join by customer id with the final customer device data
cust_sleep2 <- cust_sleep %>% 
  left_join(cust_dev3)

# removes sex categories that are empty 
cust_sleep2 <- cust_sleep2 %>% 
  filter(!is.na(sex))

# Final cust_sleep data
# Also removes users that use yellow emojis to use later for model
cust_sleep3 <- cust_sleep2 %>% 
  filter(skin_tone != "yellow")

# Displays table of mean and variance of each skin tone which will be later manually formatted in a table
cust_sleep3 %>%
  group_by(skin_tone) %>%
  summarise('Mean' = mean(flags), "Variance" = var(flags))
```

```{r}
#write data to data folder
write_rds(cust_dev3, "data/cust_dev3.Rds")

write_rds(cust_sleep3, "data/cust_sleep3.Rds")

```
